---
title: "Machine Learning PS4:ML trading Algo"
author: "YiTao Hu"
date: "26/04/2020"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
#import data and libraries
library(readr)
library(data.table)
library(foreign)
library(PerformanceAnalytics)
library(glmnet)
library(ggplot2)
StockRetAcct_DT= as.data.table(read.dta("StockRetAcct_insample.dta")) #set the key as firm ID and year
setkey(x = StockRetAcct_DT,FirmID,year)
StockRetAcct_DT=na.omit(StockRetAcct_DT)
```

###Question 1
#a
```{r}
#create the new feasures 
Features_df=StockRetAcct_DT[,list(
                                lnIssue=lnIssue,
                                lnIssue2=lnIssue^2,
                                lnIssueME=lnIssue*lnME,
                                lnProf=lnProf,
                                lnProf2=lnProf^2,
                                lnProfME=lnProf*lnME,
                                lnInv=lnInv,
                                lnInv2=lnInv^2,
                                lnInv=lnInv*lnME,
                                lnME=lnME,
                                lnME2=lnME^2
                                ),
                            by=list(FirmID,year)]

Returns=StockRetAcct_DT[,1:4,]
```

#i

Note in order to use the formula written in 48 page, we need to first normalize all our features cross-sectionally (because we assume zero mean and 1 variance.), and then using the formula to compute feature-based long-short portfolio returns. The method is similar to using Fama-MacBeth Reregression for Portfolio construction after reviewing notes of Lecture 1. The mean return of Factor portfolios are shown below.


```{r}
#define a function to normalize the features
normalize=function(vec){
  return((vec-mean(vec)))
}
FactorRet=data.frame(matrix(NA,35,12))
i=1
#cross-sectionally normalize
for (t in min(Features_df$year):max(Features_df$year)){
  Xt=as.matrix(Features_df[Features_df$year==t,3:ncol(Features_df)])
  #get excess Return
  Rt=exp(Returns[year==t,,]$lnAnnRet)-exp(Returns[year==t,,]$lnRf)
  #normalize 
  Xt=apply(Xt, 2, FUN=normalize)
  #insert 1s
  Xt=cbind(1,Xt)
  #compute factor returns
  FactorRet[i,]=t(Xt)%*%Rt/nrow(Xt)
  #FactorRet[i,]=solve(t(Xt)%*%Xt)%*%t(Xt)%*%Rt
  i=i+1
}
colnames(FactorRet)=c(colnames(Features_df[,3:ncol(Features_df),]),'EWMkt')
row.names(FactorRet)=1980:2014
colMeans(FactorRet)
```

The Var-Cov matrix is shown below

```{r}
cov(FactorRet)
```

The sample SR is shown below.
```{r}
FactorRet=xts(FactorRet,order.by=seq(from=as.Date('1980-01-01'),to=as.Date('2014-01-01'),by='years'))

SharpeRatio(FactorRet,FUN = 'StdDev')
```

##ii

Here, we perform C.V to choose the optimal lambda.
```{r}
#define a Cost function
Costfn=function(Fitted,Truths){
  return(mean((Fitted-Truths)^2))
}
TrainsetCV=FactorRet[1:25,]
Testset=FactorRet[26:35,]

#further divide the train data into training set and cv set
CV_coefs=c()

for (t in seq(1,21,by=5))
  {
  CVset=TrainsetCV[t:(t+4),]
  Trainset=TrainsetCV[-t:-(t+4),]
  Fbar_train=colMeans(Trainset)
  Fbar_cv=colMeans(CVset)
  X_train=cov(Trainset)
  X_cv=cov(CVset)
  Model=glmnet(x = X_train,y = Fbar_train,family = 'gaussian',alpha = 0.5)
  Fiited_cv=predict(Model,X_cv)
  Costs=apply(Fiited_cv, 2, FUN=Costfn,Truths=Fbar_cv)
  CV_coef=cbind(Model$lambda,Costs)
  CV_coefs=rbind(CV_coefs,CV_coef)
}
#compute mean Cost
colnames(CV_coefs)=c('Lambda','Cost')
CV_coefs=data.frame(CV_coefs)
CV_coefs=aggregate(CV_coefs,by=list(CV_coefs$Lambda),FUN=mean)
#find opt labmda
opt_lambda=CV_coefs[CV_coefs$Cost==min(CV_coefs$Cost),]$Lambda
```

Then we use optimal lambda to re-estimate the Model, the betas are shown below 
```{r}
OptModel=glmnet(x=cov(TrainsetCV),y = colMeans(TrainsetCV),family = 'gaussian',lambda = opt_lambda,intercept = F)
bs=OptModel$beta
bs
```

##iii
Finally we can compute the MVE Portfolio returns, which is given by:
$$R^{Port}_t=b^TF_t$$
```{r}
MVE_Ret=t(bs)%*%t(as.matrix(Testset))
```


Then, we can compute the out-of-sample Performance statistics
```{r}
out_sample_test=list(Mean=mean(MVE_Ret),SD=sd(MVE_Ret),SR=mean(MVE_Ret)/sd(MVE_Ret))
out_sample_test
```

##iv

Note, to be realistic, we real-estimate our model each year and compute the MVE Portfolio in accorance. 

```{r}
#define a function to map from trainCVset to coeffs
getbs=function(TrainsetCV)
  {
#further divide the train data into training set and cv set
CV_coefs=c()

for (t in seq(1,(nrow(TrainsetCV)-4),by=round(nrow(TrainsetCV)/5)))
  {
  CVset=TrainsetCV[t:(t+4),]
  Trainset=TrainsetCV[-t:-(t+4),]
  Fbar_train=colMeans(Trainset)
  Fbar_cv=colMeans(CVset)
  X_train=cov(Trainset)
  X_cv=cov(CVset)
  Model=glmnet(x = X_train,y = Fbar_train,family = 'gaussian',alpha = 0.5)
  Fiited_cv=predict(Model,X_cv)
  Costs=apply(Fiited_cv, 2, FUN=Costfn,Truths=Fbar_cv)
  CV_coef=cbind(Model$lambda,Costs)
  CV_coefs=rbind(CV_coefs,CV_coef)
}
#compute mean Cost
colnames(CV_coefs)=c('Lambda','Cost')
CV_coefs=data.frame(CV_coefs)
CV_coefs=aggregate(CV_coefs,by=list(CV_coefs$Lambda),FUN=mean)
#find opt labmda
opt_lambda=CV_coefs[CV_coefs$Cost==min(CV_coefs$Cost),]$Lambda
OptModel=glmnet(x=cov(TrainsetCV),y = colMeans(TrainsetCV),family = 'gaussian',lambda = opt_lambda,intercept = F)
bs=OptModel$beta
return(bs)
}
MVE_Port_Ret=c()
#loop through time 
for (t in 1:10){
  bs=getbs(TrainsetCV = FactorRet[1:(t+24),])
  Ret=as.numeric(t(bs)%*%t(as.matrix(Testset[t,])))
  MVE_Port_Ret=c(MVE_Port_Ret,Ret)
}
MVE_Port=cumprod(1+MVE_Port_Ret)
#compute VW port Ret
StockRetAcct_DT[,EXRet:=exp(lnAnnRet)-exp(lnRf),]
VWRET=StockRetAcct_DT[,list(VWRet=weighted.mean(x = EXRet,w = lnME)),by=year]
VWRET=VWRET[order(year)]
VWRET=VWRET[year>2004,]
VWRET[,CumRet:=cumprod(VWRet+1),]
#set the leverage factor
k=sd(VWRET$VWRet)/sd(MVE_Port_Ret)
MVE_Port_Ret=k*MVE_Port_Ret
#plot the returns
plt=ggplot(VWRET,aes(x = year,y = CumRet,color='VWMarket'))
plt+geom_line()+geom_line(aes(y=MVE_Port,color='MVEPort'))
```

